---
title: "ECO602- Using Models Activity"
author: "Mercy Melo"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(here)
catrate <- read.csv(here("data", "catrate.csv"), header = TRUE, sep = ",")
```

# Question 1
Create a histogram of the salamander reproduction catastrophic rates.

```{r catrate hist}
hist(catrate$cat.rate, 
     main = "Histogram of \n Salamander Catastrophic Rates", 
     xlab = "Catastrophic Rate")
```

# Question 2
Conduct a Shapiro-Wilk test of normality of the salamander catastrophic rates. Report the p-value and show the R-code you used to conduct the test.

p = 0.04097

```{r shapiro wilk}
shapiro.test(catrate$cat.rate)
```

# Question 3
What is the null hypothesis for the Shapiro test?

The null hypothesis of the Shapiro test is that the data are normally-distributed.

# Question 4
Based on the Shapiro test results, is there strong evidence that the sample came from a non-normally-distributed population?

Yes, the data seem to be from a non-normally-distributed population since the p value was less than the significance threshold of 0.05.

# Question 5
Show the code you used to conduct the t-test.

```{r t-test}
t.test(catrate$cat.rate, mu = 0.28)
```

# Question 6
State the null hypothesis of the test, in plain nontechnical English.

The null hypothesis is that the catastrophe rate (or the proportion of years where there are less than 1 offspring produced per breeding female) is the same as the late-fill rate (or the proportion of times the vernal pools filled later than normal).

# Question 7
 Is this a one- or two-tailed test?
 
 This is a two-tailed test because the catastrophe rate can be higher or lower than the late-fill rate.
 
# Question 8
What is the p-value from your t-test? Interpret the p-value as a false-positive rate using nontechnical English that a non-scientist would understand.

The p-value from this test was 0.01, which means that if we replicated this experiment 100 times, we'd wrongly conclude that there was a difference between the two rates once. In other words, we have a 1% chance of being incorrect when deciding that there is a difference between the catastrophe rate and the late-fill rate. 

# Question 9
What is the confidence interval for the difference between the null hypothesis and alternative hypothesis means? Did it include zero?

CI = 0.3526250 0.7261295
This confidence interval does not include zero, which indicates that we can be 95% certain that our difference in the two rates is larger than zero. 

# Question 10
Considering the results from your t-test, did you conclude that there was strong evidence to reject the null hypothesis?

Yes, the p-value (0.01) was below our alpha significance level of 0.05, indicating that there is a significant difference between the 2 rates. 

# Question 11
Show the code you used to conduct the test.

```{r wrs test}
wilcox.test(catrate$cat.rate, mu = 0.28)
```

# Question 12
Compare the p-value with the p-value you got from the t-test.

The p-value from the Wilcoxon rank sum test (0.006) was less than the p-value from the t-test (0.01).

# Question 13
Considering the results from your rank sum test, did you conclude that there was strong evidence to reject the null hypothesis?

Yes, because the p-value (0.006) was less than the alpha significance value of 0.05, we can reject the null hypothesis.

# Question 14
Compare the overall conclusions you could draw from the results of the two tests.

Both tests suggested that the catastrophe rate was significantly different from the late-fill rate.

# Question 15
Considering the numerical and graphical data exploration, which test do you think was more appropriate for these data?

The Wilcoxon rank sum test would be the more appropriate choice for this analysis because our data were not normally-distributed (a key assumption of the t-test).

# Question 16
Show the R-code you used to conduct tests of normality for the flipper lengths of Chinstrap and Adelie penguins.

```{r normality flippers}
require(palmerpenguins)
penguin_adelie = droplevels(subset(penguins, species == "Adelie"))
penguin_chinstrap = droplevels(subset(penguins, species == "Chinstrap"))

shapiro.test(penguin_adelie$flipper_length_mm)
shapiro.test(penguin_chinstrap$flipper_length_mm)
```

# Question 17
Interpret the test results. Do you conclude that the flipper lengths are normally-distributed for each species?

Yes, the flipper lengths of both penguin species are normally-distributed because the p-values from both tests were greater than the alpha significance value of 0.05.

# Question 18
Create a single figure consisting of histograms of flipper lengths of Adelie and Chinstrap penguins.

```{r hist of flippers}
par(mfrow = c(1,2))
hist(penguin_adelie$flipper_length_mm, main = "Histogram of \n Adelie Flipper Lengths", xlab = "Flipper Lengths")
hist(penguin_chinstrap$flipper_length_mm, main = "Histogram of \n Chinstrap Flipper Lengths", xlab = "Flipper Lengths")

```

# Question 19
State the alternative hypothesis of the test, in plain nontechnical English.
 
The alternative hypothesis is that the two species of penguins have different flipper lengths from one another. 

# Question 20
Include the code you used to conduct the t-test.

```{r flipper t-test}
t.test(penguin_adelie$flipper_length_mm, penguin_chinstrap$flipper_length_mm)
```
